---
title: "Developing Data Products PeerGradedAssignmentWeek4"
date: "2 June 2018"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## The goal

This application aims to quickly perfom k-mean clustering using the *iris* dataset.
The user will be able to change:

- The parameters used for the clustering
- The number of clusters desired 
- The values using a filter


## Data overview 
First let's take a look on the data.
```{r echo = TRUE}
summary(iris)
```


## K Mean custering 

Then we cqn perform the clustering
```{r echo = FALSE, message=FALSE,warning=FALSE}
library(dplyr)
```
```{r echo = TRUE}
iris %>% select(Sepal.Length, Sepal.Width) %>% #use the first two attribute by default
    kmeans(centers=3) ->        # do k-means clustering with 3 centers by default
    km                          # store result as `km`
```
Here we can see the custering compared to the actual species.

```{r echo = TRUE}
table(km$cluster, iris$Species)
```


## Plot
    
Finally we can plot the clustering.

```{r, echo=FALSE}
 palette(c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3",
                      "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))
            
            par(mar = c(5.1, 4.1, 0, 1))
            
            plot(iris %>% select(Sepal.Length, Sepal.Width),
                 col = km$cluster,
                 pch = 20, cex = 3)
            points(km$centers, pch = 4, cex = 4, lwd = 4)
```

Here  are the shiny links: 

- [Online application](https://tristanry.shinyapps.io/PeerGradedAssignmentWeek4/)
- [server.R](https://github.com/tristanry/datasciencecoursera/blob/gh-pages/9.DevelopingDataProducts/PeerGradedAssignmentWeek4/server.R)
- [ui.R](https://github.com/tristanry/datasciencecoursera/blob/gh-pages/9.DevelopingDataProducts/PeerGradedAssignmentWeek4/ui.R])
